---
title: "Penguin Size Comparisons"
author: "Krista Finlay"
date: "11/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1: T- tests 
(used to compare means between two different groups. Asking questions of do we have enough evidence to make us think that these are drawn from populations that have different means or am I going to stick with the null hypothesis that these were drawn from populations with the same mean )

Mass comparison: Between male and female Chinstrap penguins

First do some exploring in exploratory code chunk

```{r}
#create subset only for chinstrap penguins
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")

#Make some basic plots to explore normality and distribution, and get some descriptive statistics. Two types of tests of exploratory visualizations:

##Histograms (use facet plots based on groups in variable)
ggplot(data = chinstraps, aes(x = body_mass_g))+
  geom_histogram(bins = 10)+
  facet_wrap(~sex)

#two distributions that look somewhat close to bell shape, mean would be a reasonable metric of central tendency, bell shapes somewhat symmetric, central tendency of the male weight a bit higher around 4000 with the female around 3500


##Quantile- quantile plot  (explore normality- compare our sample to data that if drawn from a normal distribution. The closer to linear a qq plot gets, the closer our sample is to normally distributed)

#QQ plot (check for normality)
ggplot(data = chinstraps, aes(sample = body_mass_g))+
  geom_qq()+
  facet_wrap(~sex)

#see they are as close to normally distributed as you can hope. Both histogram and qq telling me both the distributions are pretty close to normal.

## Need to get some descriptive statistics: (uses group_by and summarize to recognize groups within a variable or variables designate and then puts functions we call in summarize (to make columns in table), condenses the calculations for each group into values and into a summary table)
chinstraps %>% 
  group_by(sex) %>% 
  summarize(mean_mass = mean(body_mass_g),
            median_mass = median(body_mass_g),
            sd_mass = sd(body_mass_g),
            sample_size = n())
  
#see the means are a bit different, median follows the mean differences, symmetric distribution, all meaningful to us.We have finished our summary statistics.
```

### comparing sizes between male and female chinstap penguins
Get vectors of body_mass_g for male and female chinstrap penguins separately (see lab week 4.

```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)


chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)

#Now have vectors to run two sample t-tests
```

###Run a couple t-tests, in a couple different ways

1. Is there a significant difference in body mass between male and female Chinstrap penguins? 
(do we have enough evidence to make us think that means difference is far enough from zero, that we think the population mean mass is different from the groups that these were drawn from- two sided t-test)

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest


# Then need to find the Cohen d Effect size (metric used ti describe the magnitude of differences between groups)
cohen.d(chinstraps_f, chinstraps_m)

```
Are male chinstrap penguins significantly larger (body mass) then female chinstrap penguins? (one sided test, only interested in tail of t- distribution' is mean male mass significantly greater

)
```{r}
t.test(chinstraps_m, chinstraps_f, alternative = "greater")
```





## One- way ANOVA
(Way to compare means across more than two groups; single variable with discrete groups within and a continued measured variable, and we want to explore differences across our different groups across factors for the mean of that other variable that we have recorded values for). 
We are looking at flipper length between the three species of penguins and only females.

###Data Exploration
1. Table of counts by penguin species and sex

```{r}
#shows the type of penguins, and how many of each sex
penguin_counts <- penguins %>% 
  count(species, sex)

penguin_counts
```


Create a subset of only female penguins

```{r}
#Explored for normallity and checked for distribution in this set

penguin_f <- penguins %>% 
  filter(sex == "female")

#Then look at the distributions (they look somewhat symmetric, bell shaped, they close to normally distributed)
ggplot(data = penguin_f, aes(x = flipper_length_mm))+
  geom_histogram(aes(fill = species))+
  facet_wrap(~species, scales = "free")


#Look at other ways to see normallity and distribution with a qq plot (then can see they are close to linear and drawn for a normally distributed population, so very close to normally distributed groups)
ggplot(data = penguin_f, aes(sample = flipper_length_mm))+
  geom_qq(aes(color = species))+
  facet_wrap(~species, scales = "free")

```

###Find Summary statistics, nice table with information (check for equal variances too)

Starting from penguins_f, create a summary table that groups by species, and returns the following:
- mean of flipper length
- sd of flipper length
- sample size
- variance "var"
- standard error = sd / sqrt (n)

```{r}
penguin_f_summary <- penguin_f %>% 
  group_by(species) %>% 
  summarize(mean_flipper_length = mean(flipper_length_mm),
            sd_flipper_length = sd(flipper_length_mm),
            sample_size = n(),
            se_flipper_length = sd(flipper_length_mm)/sqrt(n()),
            var_flipper_length = var(flipper_length_mm))

penguin_f_summary



#How to run a Levene's test for equal variance (use with caution b/c as sample size gets larger, will detect difference in variance, even if minimal and trivial) This is looking at variance in flipper length as a function of species, adn where located

leveneTest(flipper_length_mm ~ species, data = penguin_f)

```


###Make a beeswarm plot 
(with mean and plus or minus standard deviation because it can be useful to show- can be considered type of visualization that shows summary info too) 
This will include different geoms are going to look at different data frames

```{r}
#gives you actual values of your observations, and can show the greater densities of observations are. Then I am adding another geom to add some summary statistics (points of mean values), and then add wisker with errorbar of + and - sd
ggplot()+
  geom_beeswarm(data = penguin_f, aes(x = species, y = flipper_length_mm), 
                color = "gray70", 
                size = 3, 
                alpha = 0.5)+
  geom_point(data = penguin_f_summary, aes(x = species, y = mean_flipper_length),
             size = 4)+
  geom_errorbar(data = penguin_f_summary,
                aes(x = species, 
                    ymin = mean_flipper_length - sd_flipper_length,
                    ymax = mean_flipper_length + sd_flipper_length),
                width = 0.1)+
  theme_light()

# So now have actual observations plotted, and then mean and sd information. We have done a ton of exploratory data vizualizations, thought about the data
```


### One-way ANOVA

```{r}
#compare flipper length as a function of species, only female data
penguin_aov <- aov(flipper_length_mm ~ species, data = penguin_f)
summary(penguin_aov)

#Run a post-hoc Tukeys HSD on the avo name set
TukeyHSD(penguin_aov)

```
All of the p adj are below 0.5, so significant difference between mean flipper length between all the penguin species, female mean flipper differs between each species for all para wise combinations.

Conclusion: Mean flipper lengths differed significantly between Chinstrap (mean +/- se), Gentoo (mean +/- se), and Adelie (mean +/- se) penguins (one-way ANOVA with post-hoc Tukey’s HSD, F(df) = ??, p < 0.05).


Enough evidence to reject the null hypothesis, that our samples are drawn from populations with the same mean, which means that our conclusion is that there is a significant difference in mean flipper length between all female penguin species


###If trying to do anova for a very unbalanced design, very different sample sizes, groups, distribution, that you are trying a anova

###Type III Anova with the car package (will create linear model to use)

```{r}
penguin_lm <- lm(flipper_length_mm ~ species, data = penguin_f)
penguin_aov_t3 <- Anova(penguin_lm, type = "III")
penguin_aov_t3
```



